
#"Reproducible Research: Peer Assessment 1"

This is markdown document addressing peer assigment 1 from "Reproducilbe Research"

## Loading and preprocessing the data
Lets read data from the zip file
```{r data set}
dataDF<-read.csv(unz("repdata_data_activity.zip","activity.csv"), na.strings = c("","NA"))
dataDF$date<-as.Date(dataDF$date)
```


## What is mean total number of steps taken per day?
Time to calculate total number of steps taken each day
```{r aggregation, echo=FALSE}
stepsTotal<-aggregate(steps ~ date, data=dataDF, sum)
stepsTotal
```

Now, let's plot histogram of total steps by day
```{r histogramplot, echo=FALSE, message=FALSE, warning=FALSE}
library(ggplot2)
ggplot(data=stepsTotal, aes(stepsTotal$steps)) + geom_histogram(col="red", aes(fill=..count..)) + ggtitle("Histogram of total steps by day")
```

Mean of the total number of steps taken per day:
```{r mean, echo=FALSE}
mean(stepsTotal$steps)
```
Median of the total number of steps taken per day:
```{r median, echo=FALSE}
median(stepsTotal$steps)
```

## What is the average daily activity pattern?

We are going to plot time series plot with average number of steps by day.
```{r timeseries, echo=FALSE, message=FALSE, warning=FALSE}
stepsAvg<-aggregate(steps ~ interval, data=dataDF, mean)
ggplot(aes(y=steps, x=interval), data=stepsAvg)+geom_line(size=1) + geom_point() + labs(x="Interval", y="Avg Steps") + ggtitle("Average number of steps by interval")
```

Which 5-minute interval, on avarage across all the days in the dataset, contains the maximum number of steps?
```{r interval, echo=FALSE}
stepsAvg[stepsAvg$steps==max(stepsAvg$steps),1]
```


## Imputing missing values
Time to handle missing values (if I can ;) )
Calculating the total number of missing values in the dataset
```{r missing values}
sum(is.na(dataDF$steps))
```

The strategy for filling in all the missing values in the dataset. Super simple approach - I take average number of steps for the interval.
"stepsAvg" holds number of steps by interval. 
```{r steps avg, echo=FALSE}
summary(stepsAvg)
```

A new dataset is created like this:
```{r new dataset, message=FALSE, warning=FALSE}
library(plyr)
names(stepsAvg)[2]<-"stepsmean"
dataDFFixed<-join(dataDF,stepsAvg,by="interval",type="inner")
dataDFFixed$steps<-ifelse(is.na(dataDFFixed$steps),as.integer(dataDFFixed$stepsmean), dataDFFixed$steps)
dataDFFixed<-dataDFFixed[,1:3]
```
The dataset has been fixed, we will plot histogram now:
```{r histogramfixedplot, echo=FALSE, message=FALSE, warning=FALSE}
stepsTotal<-aggregate(steps ~ date, data=dataDFFixed, sum)
ggplot(data=stepsTotal, aes(stepsTotal$steps)) + geom_histogram(col="red", aes(fill=..count..)) + ggtitle("Histogram of total steps by day for fixed dataset")
```

Mean of the total number of steps taken per day for fixed dataset:
```{r mean fixed, echo=FALSE}
mean(stepsTotal$steps)
```
Median of the total number of steps taken per day for fixed dataset:
```{r median fixed, echo=FALSE}
median(stepsTotal$steps)
```

As we can see the histogram and mean/median are quite different. The average number of steps gor lower!


## Are there differences in activity patterns between weekdays and weekends?
We are going to create a new dataset that will have weekday factor.
```{r extended dataset}
dataDFFixed$weekpart<-ifelse(weekdays(dataDFFixed$date) %in% c("Saturday","Sunday"), "weekend", "weekday")
```

Panel plot with time series plots of the 5-minute interval and the average steps taken, averaged across all weekday days or weekend days
```{r weekpartplot, echo=FALSE, message=FALSE, warning=FALSE}
stepsAvg<-aggregate(steps ~ interval+weekpart, data=dataDFFixed, mean)
ggplot(aes(y=steps, x=interval), data=stepsAvg)+geom_line(size=1) + geom_point() + facet_wrap(~weekpart, nrow=2) + labs(x="Interval", y="Avg Steps") + ggtitle("Average number of steps by interval by week part")
```
```
```
```

